### 2024-04-19 paper.Rmd
First submission Cognition

---
title             : "How wise is the crowd? \n Can we infer people are accurate and competent merely because they agree with each other?"
shorttitle        : "How wise is the crowd?"

@@

## Do people infer that individuals whose answers converge tend to be right, and to be competent?

@@

In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges' who need to make numerical estimates--sometimes on factual knowledge, e.g. 'What year was the Suez Canal opened first? [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g. 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020]. To help answer these questions, participants are given estimates from others, the 'advisors'.

@@

Other studies manipulated the degree of convergence within a group of advisors. These studies present participants with a set of estimates, some of which are close to each other , while others are outliers [@harriesCombiningAdviceWeight2004; @yanivWeightingTrimmingHeuristics1997, study 3 & 4]. These studies find that participants discount outliers when aggregating estimates.

@@

## Is inferring competence from convergence justified?

@@

### Analytical argument

@@

Now, can we infer that informants belonging to the majority are more competent? To do so, we must assume some distribution of competence in the population. Using Bayes theorem, we can then show (see ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig. \@ref(fig:figure-Benoit-2)). We also find that , the more competence varies in the population, the more the degree of convergence is indicative of an informant’s competence. This reflects the fact that, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence. More generally, the fact that the focal individual is right (their ‘accuracy’) can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa) (see ESM). Figure \@ref(fig:figure-Benoit-2) provides an example with a specific categorical choice scenario, under two different population distributions of competence.

(ref:figure-Benoit-2) Average estimated competence and accuracy for an individual, depending on how many participants agree with him, in a categorical choice scenario. Here, we assume that 20 individuals answer a 5-choice question. For values higher than four, we assume that the individual is part of the majority. Below four, we assume that the focus individual is part of a minority, and that the majority size is four. Two situations are represented: one in which there is low variance in the distribution of competence in the population (orange line in the insert), and one in which there is high variance in this distribution (blue line in the insert).

```{r figure-Benoit-2, echo=FALSE, fig.cap="(ref:figure-Benoit-2)", out.width= "50%", fig.align="left"}
knitr::include_graphics("figures/figure_Benoit-2.png")
```

### Simulations

In order to better understand the influence of different parameters (e.g. degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations. In all simulations, we assume agents to be unbiased and independent in their answers, but varying in their competence.

#### Continuous case

@@

#### Categorical case

@@

## Overview of the experiments

@@

All experiments were preregistered. Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/) (<https://osf.io/6abqy/>). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the `lme4` package and its `lmer()` function. Unless mentioned otherwise, we report unstandardized model coefficients that can be interpreted in units of the scales we use for our dependent variables.

@@

The interaction design of our third experiment made the power analysis more complex and less standard than for experiments one and two. Because we could build upon data from the first experiment, we ran a power analysis by simulation. The simulation code is available on OSF, and the procedure is described in the preregistration document. The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% of the time, given an alpha threshold for significance of 0.05. Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study -- again, from the UK and via Prolific (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).

@@

In a series of experiments, we show that participants (UK) draw these inferences: when presented with more convergent answers, they tend to believe the answers are more likely to be correct, and that the individuals who made them are more likely to be competent. This is true for numerical estimates and for categorical answers. These beliefs are weakened when the individuals making the estimates are systematically biased by a common conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).

The results--both simulations and experiments--are a novel contribution to the wisdom of crowds literature. In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers. From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate. Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be sufficient in determining who tends to be an expert. This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert [@nguyenCognitiveIslandsRunaway2020; @hahnHowGoodYour2018].

@@

The current study has a number of limitations. In our simulations, we assume agents to be independent. Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency [@ladhaCondorcetJuryTheorem1992], more robust simulations would show that–while still assuming no systematic bias–our results hold even when agents influence each others' answers. If the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the experiments. Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019]. However, we do not know whether they would generalize to other cultures.

@@

The extracted data used to produce our results are available on the OSF project page (<https://osf.io/6abqy/>).

@@

The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (<https://osf.io/6abqy/>).

@@

>>>>

---
title             : "How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?"
shorttitle        : "How wise is the crowd?"

@@

# Do people infer that individuals whose answers converge tend to be right, and to be competent?

@@

In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges' who need to make numerical estimates--sometimes on factual knowledge, e.g. 'What year was the Suez Canal opened first?' [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g. 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020]. To help answer these questions, participants are given estimates from others, the 'advisors'.

@@

Other studies manipulated the degree of convergence within a group of advisors. These studies present participants with a set of estimates, some of which are close to each other, while others are outliers [@harriesCombiningAdviceWeight2004; @yanivWeightingTrimmingHeuristics1997, study 3 & 4]. These studies find that participants discount outliers when aggregating estimates.

@@

# Is inferring competence from convergence justified?

@@

## Analytical argument

@@

Now, can we infer that informants belonging to the majority are more competent? To do so, we must assume some distribution of competence in the population. Using Bayes theorem, we can then show (see ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig. \@ref(fig:figure-Benoit-2)). We also find that, the more competence varies in the population, the more the degree of convergence is indicative of an informant’s competence. For example, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence. More generally, the fact that the focal individual is right (their ‘accuracy’) can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa). Figure \@ref(fig:figure-Benoit-2) provides an example with a specific categorical choice scenario, under two different population distributions of competence.

(ref:figure-Benoit-2) Average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario. Here, we assume that 20 individuals answer a 5-choice question. For values higher than four, we assume that the focal individual is part of the majority. Below four, we assume that the focal individual is part of a minority, and that the majority size is four. Two situations are represented: one in which there is low variance in the distribution of competence in the population (orange line in the insert), and one in which there is high variance in this distribution (blue line in the insert).

```{r figure-Benoit-2, echo=FALSE, fig.cap="(ref:figure-Benoit-2)", out.width= "75%", fig.align="left"}
knitr::include_graphics("figures/figure_Benoit-2.png")
```

## Simulations

In order to better understand the influence of different parameters (e.g. degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations. In all simulations, we assume agents to be unbiased and independent in their answers, but varying in their competence. All code and data regarding the simulations can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).

### Continuous case

@@

### Categorical case

@@

# Overview of the experiments

@@

All experiments were preregistered. Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the `lme4` package and its `lmer()` function. Unless mentioned otherwise, we report unstandardized model coefficients that can be interpreted in units of the scales we use for our dependent variables.

@@

The interaction design of our third experiment made the power analysis more complex and less standard than for experiments one and two. Because we could build upon data from the first experiment, we ran a power analysis by simulation. The simulation code is available on the OSF, and the procedure is described in the preregistration document. The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% of the time, given an alpha threshold for significance of 0.05. Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study -- again, from the UK and via Prolific (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).

@@

In a series of experiments, we show that participants (UK) draw these inferences: when presented with more convergent answers, they tend to believe the answers are more likely to be correct, and that the individuals who made them are more likely to be competent. This is true for numerical estimates and for categorical answers. These beliefs are weakened when the individuals making the estimates are systematically biased by a shared conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).

The results--both simulations and experiments--are a novel contribution to the wisdom of crowds literature. In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers. From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate. Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert. This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert [@nguyenCognitiveIslandsRunaway2020; @hahnHowGoodYour2018].

@@

The current study has a number of limitations. In our simulations, we assume agents to be independent. Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency [@ladhaCondorcetJuryTheorem1992], more robust simulations would show that–while still assuming no systematic bias–our results hold even when agents influence each others' answers. Regarding our experiments, if the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the results. Although the main results replicate well across our experiments, and we can thus be reasonably certain of their robustness, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019]. However, we do not know whether they would generalize to other cultures.

@@

Data for all experiments and the simulations is available on the OSF project page (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).

@@

The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).

@@

\clearpage
\newpage
